---
title: "Test Markdown"
author: "Becca Selden"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

### Objective
Work through RMarkdown commands with students
Last edited: November 12, 2024

# Big Header

###### Small Header

### This is a code chunk
This uses 3 backwards apostrophes (on top left of keyboard) 

```{r code chunk1}
a <- 42

#print the output
a
```

### Load a R dataset on cars
```{r code chunk 2}
mtcars
```
### Install packages
We want to install tidyverse and here as two packages
```{r install packages}
# install.packages(c("tidyverse", "here"))
# install.packages(c("lubridate"))

```

### Loading data from a url
```{r load data csv}
monster_movie_genres <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movie_genres.csv')

monster_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movies.csv')


```
`package::function()` is a way to make sure that R is using the "right" function (or the one you want it to use) if there are more than one with same name.
Read data in by assigning it to a new object (monster_movie_genres) with the backwards arrow `<-` and using read_csv() as the function to do that (in this case from a url, but could also be a csv in the Data folder

### Write a csv output
`write_csv(object_name, filename_for_saving)`
```{r write csv}
write_csv(monster_movie_genres, here::here("Data/monster_movie_genres.csv"))
write_csv(monster_movies, here::here("Data/monster_movies.csv"))

```

### Read the csv from a file on your computer
```{r read csv file}
monster_movie_genres2 <- read_csv(here::here("Data/monster_movie_genres.csv"))
```

### Merge two files together
The movie genres file just had codes for the movie
The movies file had info on the movie (title, date)
We want them together so we can look at monster movies in a genre by year
`new_data <- inner_join(data1, data2, by=c(x=y))`
```{r inner join}
monster <- inner_join(monster_movie_genres, monster_movies, by="tconst")
```

### number of genres per movie
Tidyverse syntax `%>%` = take this then do this
new_object <- old_object %>%
do_something_to_old_object
```{r num genres per movie by year}
num_genres_by_movie <- monster_movie_genres %>%
  group_by(tconst) %>%
  summarize(num_genres=n_distinct(genres))

monster2 <- inner_join(num_genres_by_movie, monster_movies, by="tconst")

```